The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-07-01 Paper/jaccardbaetica.tex
add new changes correction xavi

The landscape of the \textit{Baetican} province was one of the best regions to face the increase in demand of olive oil across the Roman provinces. For this reason the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed. The production and distribution of olive oil grew exponentially\xavi{seguro que exponencial?} for almost three centuries\xavi{qué centuries?} \citep{remesal_concierto}. Hundreds of presses and amphora-making workshops were build near large extensions covered by olive trees. The workshops where amphorae were made and filled with olive oil were places along the rivers Guadalquivir and Genil riverbanks (see Fig.\ref{workshop}).

@@

The evidence for olive oil consumption in \textit{Britannia} is scarce before the Roman conquest \citep{funari_corpus_1996,carreras_abastecimiento_2003}. Indigenous\xavi{no sé si indigenous es la mejor palabra...local?} population did not consume this product and it was not produced on the region as the landscape and climate of the British isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}. Based on the archaeological record this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} \citep[1]{carreras_britannia_1998}\xavi{en qué siglo?}. The demand of olive oil after the Roman conquest required of a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.

@@

The presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{carreras_britannia_1998}.\xavi{si dices widely studied no puedes referencias solo 1 autor}. This fact\xavi{que fact? Que los sellos de dressel 20 han sido muy estudiados? No tengas miedo a repetir trozos rollo "This intense consumption of olive oil also indicatas blablabla...} also indicates a possible provincial structure designed to organize the supply of olive oil to military camps,m such as \textit{Germania} (Remesal Rodrıguez, 1986)\xavi{por qué such as Germania? Aún no has explicado qué hacen en Germania}. There are no written records explaining how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep[45]{funari_economic_2005}.

@@

%\xavi{si el eje es logarítmico deberías mencionarlo}

@@

%\xavi{vale pero qué es cada punto? un workshop?}

@@

%\xavi{Esta estructura no me convence...tenemos que pensar algo mejor, porque ahora es: 1. hablo de britania/germania, 2. hablo de sellos en general, 3. hablo de sellos en britania/germania. Quizás juntar esto con lo de los casos de estudio, o quizás hablar de los 2 casos de estudio en la intro, y dividir la sección anterior en 2 (lo más general a la intro, lo más concreto aquí)}
The analysis of stamps in consumption areas also used the CEIPAC database and filtered in the same way than the previous dataset. We selected the centres with more or equal than five stamps\xavi{pero esto no se hizo en el de Baetica no? Por qué?}.

@@

Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}. Morisita-Horn index gives a value between 0 (sites have an exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes). To apply Morisita-Horn we firstly calculated the number of times that each stamp code appeared in an amphora workshop. This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}. If two workshops hhad similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.

@@

%\xavi{no tengo claro que le puedas decir correlación, porque eso es un valor específico entre 2 distribuciones numéricas. Te refieres a similaridad?}

@@

This trend is aligned with recent works that have suggested that the Atlantic trade routes were mostly used for military supply~\citep{remesal_annona_1986,remesal_provincial_2008,carreras_atlantic_2012,morillo_hispania_2016,rubio-campillo_provincias_2018}. Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting thast Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,ayllon_olive_2018}.

@@

%In general, we observed a concentration of the stamps in military centres.

@@

Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}. The stamped amphorae could have allowed potters to organise the differnt production batches for posterior commercialisation\xavi{estos 2 párrafos igual deberían ir a la discusión no? La conclusión debería ser algo más general y aquí estás aún interpretando resultados}.

>>>>

The landscape of the \textit{Baetican} province was one of the best regions to face the increase in demand of olive oil across the Roman provinces. For this reason the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed. The production and distribution of olive oil increased from the first to the third centuries AD ~\citep{remesal_concierto}. Hundreds of presses and amphora-making workshops were build near large extensions covered by olive trees. The workshops where amphorae were made and filled with olive oil were places along the rivers Guadalquivir and Genil riverbanks (see Fig.\ref{workshop}).

@@

The evidence for olive oil consumption in \textit{Britannia} is scarce before the Roman conquest \citep{funari_corpus_1996,carreras_abastecimiento_2003}. Local population did not consume this product and it was not produced on the region as the landscape and climate of the British isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}. Based on the archaeological record this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} \citep[1]{carreras_britannia_1998}\xavi{en qué siglo?}. The demand of olive oil after the Roman conquest required of a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.

@@

The presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{carreras_britannia_1998}.\xavi{si dices widely studied no puedes referencias solo 1 autor}.

This intense consumption of olive oil also indicates a possible provincial structure designed to organize the supply of olive oil to military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986)\xavi{por qué such as Germania? Aún no has explicado qué hacen en Germania}. There are no written records explaining how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep[45]{funari_economic_2005}.

@@

@@

@@

The analysis of stamps in consumption areas also used the CEIPAC database and filtered in the same way than the previous dataset. We selected the centres with more or equal than five stamps\xavi{pero esto no se hizo en el de Baetica no? Por qué?}.

@@

Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}. Morisita-Horn index gives a value between 0 (sites have an exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes). To apply Morisita-Horn we firstly calculated the number of times that each stamp code appeared in an amphora workshop. This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}. If two workshops had similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.

@@

@@

This trend is aligned with recent works that have suggested that the Atlantic trade routes were mostly used for military supply~\citep{remesal_annona_1986,remesal_provincial_2008,carreras_atlantic_2012,morillo_hispania_2016,rubio-campillo_provincias_2018}. Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting that Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,ayllon_olive_2018}.

@@

@@

Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}. The stamped amphorae could have allowed potters to organise the different production batches for posterior commercialisation\xavi{estos 2 párrafos igual deberían ir a la discusión no? La conclusión debería ser algo más general y aquí estás aún interpretando resultados}.
